Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-01-07 papers/wdmerger_I/paper.tex

At the start of any full simulation, we generate initial model white dwarfs by integrating the equation of hydrostatic equilibrium, taking the temperature and composition to be constant, and using the general stellar equation of state.  This results in a single non-linear equation to find the density in a zone given the conditions in the zone beneath it:
\begin{equation}
\frac{p_{i+1} - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}.
\end{equation}
This equation is a function of $\rho_{i+1}$ only since the pressure is uniquely determined by the density in this case. $\rho_i and p_i$ are known,  $g_{i+1/2}$ is the gravitational acceleration at the interface between zones $i$ and $i+1$, found by simply adding up all the mass from zones $1, \ldots, i$ to get the enclosed mass, $M_{i+1/2}$, and then setting $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$. We solve this equation for $\rho_{i+1}$ using a Newton-Raphson iteration.

>>>>

At the start of any full simulation, we generate initial model white dwarfs by integrating the equation of hydrostatic equilibrium, taking the temperature and composition to be constant, and using the general stellar equation of state.  This results in a single non-linear equation to find the density in a zone given the conditions in the zone beneath it:
\begin{equation}
\frac{p_{i+1} - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}.
\end{equation}
This equation is a function of $\rho_{i+1}$ only since the pressure is uniquely determined by the density in this case. $\rho_i$ and $p_i$ are known,  $g_{i+1/2}$ is the gravitational acceleration at the interface between zones $i$ and $i+1$, found by simply adding up all the mass from zones $1, \ldots, i$ to get the enclosed mass, $M_{i+1/2}$, and then setting $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$. We solve this equation for $\rho_{i+1}$ using a Newton-Raphson iteration.
